---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---
---
title: "Final Project"
author: "Neil Farrugia 17336831"
date: "14/11/2021"
output:
  pdf_document: default
  html_document: default
---



<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Rcpp)
library(inline)
path <- "~/Masters/CProg/Assig2/RanWalk"
setwd(path)
sourceCpp("ranwalk.cpp")
```

## Question 1

I have coded up 2 walk functions. Their differences will be further explained in Question 3.
However for the assigment walk2 function will be used.\
Both these functions output a matrix of the movments as shown below:\
```{r}
print("1st function")
walk(10,seed=3)
print("2nd function")
walk2(10,seed=3)

```

\
To make sure function works I've codeed up a quick test function, that works out the probabilities of each movement happening. If it is roughly 0.25 across all 4 possibilities then we know that the algorithm is correct. Small differences may occur but it's the big differences in probaility that we would be worried about.\
```{r}
res2 <- matrix(rep(NA,1000*4), nrow = 100, ncol = 4)
res <- matrix(rep(NA,1000*4), nrow = 100, ncol = 4)
for(i in 1:100){
  
  res2[i,]<-test(walk2(10000,i))
  res[i,]<-test(walk(10000,i))
  
}
apply(res2,2,mean)
apply(res,2,mean)

```
\
Seems like function roughly have 0.25 probabilities across all 4 movements options.\
\
\
Has the destination [1,-3] been reached after walking a 100 roads? \
For this I've written a second function called Destination. In this function it calls the walk2 function and then checks whether [1,-3] has been reached. \
It returns a Boolean vector. Although the question specifically asks for TRUE or FALSE output. We can wrap it in an R function and check if there's a TRUE in the vector. For me this makes more sense as you can then check whether the destination was reached more than once and you can check after how many moves the destination was reached.\
\
```{r}
TRUE%in%Destination(as.integer(c(-1,3)),100,seed = 1) #yes destination has been reached
which(Destination(as.integer(c(-1,3)),100,seed = 1)==TRUE) #after how many movements has the destination been reached

```


# Question 2
\
Checking the probability of reaching the destination within 20 roads taken.
```{r}
yes20 = rep(0,1000) #stores 1 if reached within 20 and 0 if not
for(i in 0:1000){
  walks = Destination(as.integer(c(1,-3)),20,seed = i)
  #
  if(TRUE%in%walks){
    yes20[i] = 1
  
  }
}
sum(yes20)/1000#sum the 1s and dividde by 1000 to get mthe probaility
```
\
\
# Question 3
\
\
Originally I had only written one random walk function called walk.\
It worked as follow:\
The function takes an input up N (the number of “walks/movement” taken by the walker) and an integer for seeding.\
- declare a matrix of N rows and 2 columns\
- column 0 is the x axis and column 1 is the y axis\
- the rows are each movement\
- row index 0 is [0,0] starting position\
- the for loop starts here, from index 1 to N\
- Here there's the first coin flip, if lands on heads the walker will walk somewhere along the x axis, tails would be y axis.\
- Then another coin is flipped, if heads then walker will move forward by one, tails would be move back by one.\
- The result of these to IF-THEN statements would be with either [0,+1] , [0,-1], [+1,0],[-1,0]\
- two 50% chances = 25% chance of getting one of those options.\
- Then the movement is added to the previous position, in short position[i] = position[i-1] +movement [i]\
- The output to the function is a matrix of all the positions from index 0 to index N-1\
This for me made the most sense in my head, pick the direction then pick forward or back movement.\
\
However after reading question 3 (after I’ve written the code), I wondered is this the most efficient way of writing the algorithm. Therefore I wrote a second walk function called… walk2. And kept the first walk function to compare speed. I wanted to write the fastest possible algorithm.\
\
\
In walk2 I use a switch statement instead of an IF-ELSE statement. Switch statments are actually faster than IF-ELSE statements so the code becomes already more efficient as there’s one switch statement in comparison to 4 IF-ELSE statements, it’s also easier to follow.\
\
The next big change I made on walk2 is the implementation of bit manipulation or bit fiddling. Cpp allows you to to directly use bit operations on 32bit integers such as the integer that rand() produces.\
More precisely the code runs as follows:\
\
- There is two loops, the outer loop contains the rand() function and the bit operations.\
- The inner loop contains the switch statements and thus the movement addition part of the algorithm.\
- When rand() is called, a 32 bit integer is stored. And thus a random “string” of 1s and 0s are stored. So why not use the 16 pairs of random 1s and 0s which would produce 16 random movements rather than the actual number that is stored which would produce 1 random movement.\
- The first bit operation is “>>” which is a right shift. This shifts the 1s0s to the right each time by a factor of 2. Then the “&” operation is used, which is also called bit masking. It basically makes every 1s and 0s turn to 0 apart from the last two bits which stays the same. Meaning that at every loop we are only interested in the last 2 bits. Which can either be, 00,11,01 or 10. Thus creating 25% chances, which is the probability that the algorithm needs.\
- Next in the inner loop the switch cases are in hexagon decimal, ie 0x00,0x01,0x02 or 0x03 which corresponds to the 4 possible last two bits the integer can hold.\
- What all this means in effect is that rand() which is a costly function for the computer but necessary for the algorithm is called every 16 times. This is again because in the 32 bit integer there 16 pairs of bits, so at every loop in the inner loop, the randomness is created by each pairs of 0s and 1s in the 32bit  number\
- This makes the algorithm dramatically quicker. It also means that this algorithm more scalable because it becomes more efficient relative to the first algorithm as the number of roads/decisions increases.\
- However
\
Exactly how more efficient, can be shown with the code below;:\
\

```{r}
#----------
#compare the two walk functions
#library(rbenchmark)
#benchmark(walk(100,2),walk2(100,2),replications(100),order = "relative") 
#error using benchmark error so it wont work - "Error in formula.default(object, env = baseenv()) : invalid formula"
#-------
#Use another package instead called microbenchmark
require(microbenchmark)
microbenchmark(walk = {walk(1000000,2)},walk2 = {walk2(1000000,2)}, times = 100, unit = "milliseconds")

```
\
As you can see the walk2 is much quicker in comparison to the first walk function\
\
Also worth mentioning that I've coded up the specific functions that answers question 1,2 and another function that answers question 4. This of course is very slightly less efficient because one extra loop has to be run through the matrix produced by the walk2 function in order to answer the questions. Instead of let's say doing everything in the 1 loop in the walk2 function.\
\
The added benefit of separating out the function are:
- makes the code more readable as there is way less code.
- it makes it easier to debug if something has gone wrong in either the algorithm or the function that answers the questions\
- it also opens the door for more functions ot be made that builds on the walk2 algorithm. Such as a plotting function.\
\


# Question 4

```{r}
#Manhattan distance
avgD <- c()
for(i in 1:1000){
  avgD<-c(avgD,ManhattanD(50,i))
}
summary(avgD)
mean(avgD)

```







